---
title: "Statistical Inference Course Project. Simulation Exercise"
author: "Mariia Danilenko"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction  
In this project we will make some research on the exponential distribution in R and compare it with the Central Limit Theorem. We will investigate the distribution of averages of 40 exponentials. We will need to do a thousand simulations.  
Here and further fix $\lambda=0.2$
```{r, echo=FALSE}
lambda<-0.2
exp_means = NULL
exp_sd = NULL
expDistSimul <- matrix(1:40000, nrow=1000)
```

# Analysis    
Let's simulate 1000 averages of 40 exponential iid values.  
Also, calculate mean for each of them.
```{r}
 n<-1000
for (i in 1:n) expDistSimul[i,]<-rexp(40,lambda)
for (j in 1:n) exp_means[j] <- mean(expDistSimul[j,])
```
Based on the properties of the mean theoretical distribusion mean would be:
$$E(E(\frac{1}{40}\sum^{40}_{i=1}\chi_{exp,i}))=E(\frac{1}{40}\sum^{40}_{i=1 }E(\chi_{exp,i}))=E(\frac{1}{40}*40*\frac{1}{\lambda})=\frac{1}{\lambda}$$
In this case mean equals 5. Compare simulation mean with theoretical:

```{r}
t.test(exp_means, alternative = "two.sided", mu=1/lambda)
```
We failed to reject our hypothesis about mean.  
Speaking about sd, we apply **Lindeberg–Lévy CLT** to our model:
$$\sqrt {40} ((\frac{1}{40}\sum^{40}_{i=1}\chi_{exp,i})-\frac{1}{\lambda}) \to N(0,\sigma^2)$$
$$\sqrt {40} *(mean(rexp(40,\lambda))-\frac{1}{\lambda}) \to N(0, \frac{1}{\lambda}) \Rightarrow sd(mean(rexp(40,\lambda)) \to \frac{1}{\lambda \sqrt{40}}=0.79$$
We want to check variances for hypothesis that final distribution almost $N(\frac{1}{\lambda},\frac{1}{\lambda \sqrt{40}})=N(5,0.79)$. Let's check variance:
```{r}
 var.test(exp_means,rnorm(1000,5,0.79))
```
We failed to reject out hypothesis.  
Let's plot our distribution and 1000 instances $~N(5,0.79)$.
```{r}
par(mfcol=c(2,1))
hist(exp_means, col="lightblue",xlim = c(3,7), 
     xlab="Value", main = "Histogram of current model")
hist(rnorm(1000,5,0.79),xlim = c(3,7), col="lightgreen", 
     xlab="Value", main = "Histogram of normal distribution")
```

# Conclusion  
Our investigation coincides with CLT and final distribution is $~N(5,0.79)$
