---
title: "Basic Inferential Data Analysis"
author: "Mariia Danilenko"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction  
We're going to analyze the ToothGrowth data in the R datasets package. For this project we only use the techniques from class, even if there's other approaches worth considering.  

# Analysis    

## Given data

 ***ToothGrowth data***:  
 
The response is the length (**len** column) of odontoblasts (cells responsible for tooth growth) in **60** guinea pigs. Each animal received one of three dose (**dose** column) levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods: orange juice (OJ) or ascorbic acid (VC) (**supp** column).
Take a look to observing data:
```{r}
str(ToothGrowth)
```

##Visual analysis

Make an introduction plot of length factored by supplement type and depends on dose:  
```{r, message=FALSE, warning=FALSE, fig.height=3}
library(lattice)
par(mfcol=c(2,1))
xyplot(len ~ dose | factor(supp,labels=c("Orange juice","Ascorbic acid")),
       data = ToothGrowth, type=c("p","smooth"),
       xlab = "Dose", col.line="red",
       ylab="Tooth length")
```
Notice correlation between dose and length. Before we begin analysis, let's plot boxplot factored by dose:  
```{r plot, fig.height=3}
with(ToothGrowth, 
     boxplot(len~dose, xlab="Dose", ylab="Tooth length", 
             col=c("orange","yellow", "lightgreen")))
```

## Calculations
Cut given data into six groups: 2 sup types and 3 dose types:  
```{r}
TGOJ0.5<-subset(ToothGrowth, ToothGrowth$supp=="OJ" & ToothGrowth$dose==0.5)
TGVC0.5<-subset(ToothGrowth, ToothGrowth$supp=="VC" & ToothGrowth$dose==0.5)
TGVC1<-subset(ToothGrowth, ToothGrowth$supp=="VC" & ToothGrowth$dose==1)
TGOJ1<-subset(ToothGrowth, ToothGrowth$supp=="OJ" & ToothGrowth$dose==1)
TGOJ2<-subset(ToothGrowth, ToothGrowth$supp=="OJ" & ToothGrowth$dose==2)
TGVC2<-subset(ToothGrowth, ToothGrowth$supp=="VC" & ToothGrowth$dose==2)
```

And also select two subsets with orange juice and ascorbic acid:  
```{r}
TGOJ<-subset(ToothGrowth, ToothGrowth$supp=="OJ")
TGVC<-subset(ToothGrowth, ToothGrowth$supp=="VC")
```

### Comparisons of doses  
Comparison between 0.5 and 1 doses
```{r}
t.test(cbind(TGOJ0.5$len,TGVC0.5$len), cbind(TGOJ1$len,TGVC1$len), 
       alternative = "less")$conf.int
```
We reject $H_0$ hypothesis and claim that length mean for 0.5 dose is less than length mean for 1 dose.
Comparison between 1 and 2 doses
```{r}
t.test(cbind(TGOJ1$len,TGVC1$len), cbind(TGOJ2$len,TGVC2$len), 
       alternative = "less")$conf.int
```
We reject $H_0$ hypothesis and claim that length mean for 1 dose is less than length mean for 2 dose.  

We don't need compare 0.5 and 2 doses because of transitivity.  

Also, we checked same hypothesis for different sups separately and got same results. R code for these calculations you can find in the appendix.
Finally we claim that dose changes are significant for tooth growing.   

### Comparisons of sups  

We compared two sup types for every dose:  

```{r}
t.test(TGOJ0.5$len,TGVC0.5$len, alternative = "greater")$conf.int
t.test(TGOJ1$len,TGVC1$len, alternative = "greater")$conf.int
t.test(TGOJ2$len,TGVC2$len, alternative = "greater")$conf.int
```
We reject $H_0$ hypothesis in first and second models. It means that orange juice is more effective than ascorbic acid on 0.5 and 1 doses.  
We fail to reject $H_0$ hypothesis in fhird model, however we can explain it:  
Exist maximum level of vitamin C. which pigs can absorb. That's why in 2 dose sup type doesn't matter.
   
# Results  
We made following conclusions:  
 * dose growing is significant for tooth growing;    
 * sup type impact on tooth length until 2 dose.  

# Appendix  
Doses hypothesis separately for sups:
```{r}
# for orange juice
t.test(TGOJ0.5$len,TGOJ1$len, alternative = "less")$conf.int
t.test(TGOJ1$len,TGOJ2$len, alternative = "less")$conf.int
# for ascorbic acid
t.test(TGVC0.5$len,TGVC1$len, alternative = "less")$conf.int
t.test(TGVC1$len,TGVC2$len, alternative = "less")$conf.int
```
We reject $H_0$ hypothesis for all 4 models.


Sup whole hypothesis:  
```{r}
t.test(TGOJ$len,TGVC$len, alternative = "greater")$conf.int
```
We reject $H_0$ hypothesis.

